\documentclass[12pt, oneside]{book}
\usepackage{amsmath, amssymb, amssymb, mathtools, graphicx, hyperref, titling}
\usepackage{tikz, caption, subcaption, enumitem, polyglossia}
\usepackage{tikz-3dplot, float, etoolbox, enumitem}
\usepackage[top=25mm, bottom=25mm, left=20mm, right=20mm]{geometry}
%\usepackage{showframe}

\makeatletter
\patchcmd{\@makechapterhead}{\vspace*{50\p@}}{}{}{}% Removes space above \chapter head
\patchcmd{\@makeschapterhead}{\vspace*{50\p@}}{}{}{}% Removes space above \chapter* head
\makeatother

\predate{}
\postdate{}
\date{}
\title{Rings and Modules}
\author{Nachiketa Kulkarni}
\pagenumbering{gobble}
\setmainfont{Comic Sans MS}

\begin{document}
\maketitle
\tableofcontents

\mainmatter
\chapter{Introduction to Rings}
\section{Definition of a Ring}
A ring \(R\) is a set with two binary operations, \(+\) and \(\times\), satisfying the following conditions:
\begin{itemize}
    \item \(\left(R,+\right)\) is an abelian group.
    \item \(\times\) is associative.
    \item \(\times\) distributes over \(+\).
\end{itemize}
A Ring is said to be commutative if \(a \times b = b \times a\) for all \(a, b \in R\).
A Ring is said to have a multiplicative identity if there exists an element \(1 \in R\) such that \(1 \times a = a \times 1 = a\) for all \(a \in R\).

\paragraph{Subrings:} A subset \(S\) of a ring \(R\) is called a subring if:
\begin{itemize}
    \item \(S\) is closed under addition and multiplication.
    \item \(S\) contains the additive identity \(0\) of \(R\).
    \item For every \(a \in S\), \(-a \in S\).
\end{itemize}
\subsection{Examples}
\begin{itemize}
    \item \textbf{Trivial Ring:} Take any abelian group \((G,+)\) and define multiplication as \(a \times b = 0\) for all \(a, b \in G\), where \(0\) is the identity of the group.
    \item \textbf{Integers:} The set of integers \(\mathbb{Z}\) with usual addition and multiplication forms a ring.
    Also, the quotient group \(\mathbb{Z}/n\mathbb{Z}\) is a ring for any integer \(n\).
    \item \textbf{Hamiltonian Quaternions:} The set of quaternions \(\mathbb{H} = {1, i, j, k}\), where \(i^2 = j^2 = k^2 = -1\).
    \item \textbf{Polynomial Rings:} Fix a commutative ring \(R\).
    The set of polynomials with coefficients in \(R\), denoted \(R[x]\), forms a ring with addition and multiplication defined as usual.
\end{itemize}
\section{Properties of Rings}
\paragraph{Proposition:} If \(R\) is a ring, then the following hold:
\begin{enumerate}
    \item \(0a = a0 = 0\) for all \(a \in R\).
    \item \((-a)b = a(-b) = -(ab)\) for all \(a, b \in R\).
    \item If the ring has a multiplicative identity \(1\), then it is unique.
    \item \((-1)a = -a\) for all \(a \in R\).
\end{enumerate}
\paragraph{More Definitions:} Consider a ring \(R\):
\begin{itemize}
    \item A non-zero element \(a \in R\) is called a \textbf{zero divisor} if there exists a non-zero \(b \in R\) such that either \(ab = 0\) or \(ba = 0\).
    \item Assume \(R\) has a multiplicative identity \(1\).
    An element \(a \in R\) is called a \textbf{unit} if there exists an element \(b \in R\) such that \(ab = ba = 1\).
    The set of all units in \(R\) is denoted by \(R^\times\).
    \item A Ring \(R\) with identity is called an \textbf{integral domain} if it has no zero divisors and \(1 \neq 0\).
\end{itemize}
\paragraph{Proposition:} If \(R\) is an integral domain, then the following hold:
\begin{enumerate}
    \item \(R^\times\) is a group under multiplication.
    \item \(R\) is a field if multiplication is commutative and every non-zero element is a unit, i.e., \(R^\times = R - \{0\}\).
    \item A zero divisor cannot be a unit and vice versa.
    \paragraph{Proof:} If \(a\) is a zero divisor, then there exists a non-zero \(b\) such that \(ab = 0\).
    Now, assume \(a\) is a unit, then there exists \(c\) such that \(ac = 1\).
    But:
    \[b = (ca)b = c(ab) = c0 = 0\]
\end{enumerate}
\section{Homomorphisms and Isomorphisms}
Let \(R\) and \(S\) be rings.
A \textbf{ring homomorphism} is a function \(\phi: R \to S\) such that:
\begin{enumerate}
    \item The map \(\phi\) preserves addition: \(\phi(a + b) = \phi(a) + \phi(b)\) for all \(a, b \in R\).
    \item The map \(\phi\) preserves multiplication: \(\phi(ab) = \phi(a) + \phi(b)\) for all \(a, b \in R\).
\end{enumerate}
The kernel of a ring homomorphism \(\phi\), \(\ker (\phi)\), is the set of elements in \(R\) that map to \(0\) in \(S\).
The Image of a ring homomorphism \(\phi\), \(\text{Im}(\phi)\), is the set of elements in \(S\) that are images of elements in \(R\).
A bijective ring homomorphism is called a \textbf{ring isomorphism}, denoted by \(R \cong S\).

\subsection{Properties of Ring Homomorphisms}
Let \(R\) and \(S\) be rings and \(\phi: R \to S\) be a ring homomorphism.
Image of \(\phi\) is denoted by \(\text{Im}(\phi)\) and kernel of \(\phi\) is denoted by \(\ker\).
\paragraph{Proposition:} \(\text{Im}(\phi)\) is a subring of \(S\).
\paragraph{Proof:} \(\text{Im}(\phi)\) is a subring of \(S\) because:
\begin{itemize}
    \item \textbf{Closure under addition:} If \(x, y \in \text{Im}(\phi)\), then there exist \(a, b \in R\) such that \(\phi(a) = x\) and \(\phi(b) = y\).
    Now, \(\phi(a + b) = \phi(a) + \phi(b) = x + y\), hence \(x + y \in \text{Im}(\phi)\).
    \item \textbf{Closure under multiplication:} If \(x, y \in \text{Im}(\phi)\), then there exist \(a, b \in R\) such that \(\phi(a) = x\) and \(\phi(b) = y\).
    Now, \(\phi(ab) = \phi(a)\phi(b) = xy\), hence \(xy \in \text{Im}(\phi)\).
    \item \textbf{Associativity of Addition and Multiplication} Inherited from the ring.
    \item \textbf{Additive Identity} \(\phi(0) = 0\)
\end{itemize}
Hence, \(\text{Im}(\phi)\) is a subring of \(S\).
\paragraph{Proposition:} \(\ker(\phi)\) is a subring of \(R\).
Also, if \(\alpha \in R\), then \(\{r \alpha, \alpha r\} \in \ker(\phi)\), \(\forall r \in R\).
\paragraph{Proof:} Part 1 of the proof is same as above.
For the second part, let \(\phi(\alpha) = 0\) and \(\phi(r) = a\).
\[0 = 0 a = \phi(\alpha)\phi(r) = \phi(\alpha r) \hspace{50pt} 0 = a0 = \phi(r) \phi(\alpha) = \phi(r \alpha)\]

\subsection{Ideals}
\paragraph{Definition:} Let \(R\) be a ring, \(I\) be a subgroup of \(R\). Let \(r \in R\):
\begin{enumerate}
    \item \(rI = \{ra \; \vline \; a \in R\}\) and \(Ir = \{ar \; \vline \; a \in R\}\) 
    \item A subgroup \(I\) is called a left Ideal of \(R\) if:
    \begin{itemize}
        \item \(I\) is a subring of \(R\).
        \item \(I\) is closed under left multiplication by elements from \(R\), i.e., \(rI \subseteq I\)
    \end{itemize}
    The right Ideal is similarly defined.
    \item If \(I\) is a botha left Ideal and right Ideal, then it is called an Ideal (two sided) of \(R\).
\end{enumerate}

\subsection{First Homomorphism Theorem}
\paragraph{Theorem:}
\begin{enumerate}
    \item If \(\phi: R \rightarrow S\) is a ring homomorphism, then \(\ker(\phi)\) is an ideal of \(R\) and \(R/\ker(\phi) \cong \phi(R)\).
    \item If \(I\) is an ideal of \(R\):
    \[R \rightarrow R/I \hspace{15pt} \text{defined by} \hspace{15pt} r \rightarrowtail r + I\]
    is a surjective ring homomorphism with the kernel being \(I\).
    Thus every ideal is the kernel of a ring homomorphism and vice-versa.
    This above homorphism is known as Natural Projection of \(R\) onto \(R/I\) 
\end{enumerate}
\paragraph{Proof:}
\end{document}
